Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems\thanks{This research was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-10-18 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
phi -> f
fix remark

\textbf{Rule 4.} In the criteria space on the base of the marked up matrix $A^i_p$, a separating hyperplane $L_p$ is constructed applying logistic regression \cite{Yu2011} as efficient tool of machine learning.

\textbf{Rule 5.} For each point $x_j, 1 \leq j \leq p$, the distance $d_j$ of the vector $(\varphi_1(x_j), \dots, \varphi_k(x_j))$ to the separating plane $L_p$ is determined. This distance is considered to be negative if vector is placed above the plane and positive otherwise (see Fig. \ref{fig:1}).

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{fig1}
\caption{An example of estimating distances to the separating plane $L_p$.} \label{fig:1}
\end{figure}

@@

\textbf{Remark}. The above method is based on the ISA  global search algorithm, which has an explicit geometric interpretation (see \cite{Strongin2000}). Using trial points, it is possible to design a minorant for the objective function $f(x)$. Then the value of characteristic, taken with the opposite sign, i.e. $-R(i)$, is the minimum value of the minorant in the interval $(x_{i-1}, x_i)$. Moreover, the point of a new trial $x^{p+1}$, computed according to (\ref{eq:26}), is carried out at the point, at which the minimum of the minorant is situated.

>>>>

\textbf{Rule 4.} A hyperplane $\Pi_p$ is constructed that separates the point classes in the space of criteria labeled according to Rule 3. The classification algorithm based on logistic regression is used as a specific method of hyperplane construction \cite{Yu2011}. Note that these two classes are unbalanced in terms of the number of points, therefore, when building a hyperplane, non-dominated points are taken into account with the weight 0.95, and dominated ones with the weight 0.05.

\textbf{Rule 5.} For each point $x_j, 1 \leq j \leq p$, the distance $d_j$ of the vector $(\varphi_1(x_j), \dots, \varphi_k(x_j))$ to the separating plane $\Pi_p$ is determined. This distance is considered to be negative if vector is placed above the plane and positive otherwise (see Fig. \ref{fig:1}).

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{fig1}
\caption{An example of estimating distances to the separating plane $\Pi_p$.} \label{fig:1}
\end{figure}

@@

\textbf{Remark}. The above method is based on the ISA  global search algorithm, which has an explicit geometric interpretation (see \cite{Strongin2000}). Using trial points, it is possible to design a minorant for the objective function $f(x)$. Then the value of characteristic, taken with the opposite sign, i.e. $-B(i)$, is the minimum value of the minorant in the interval $(x_{i-1}, x_i)$. Moreover, the point of a new trial $x^{p+1}$, computed according to (\ref{eq:26}), is carried out at the point, at which the minimum of the minorant is situated.
